---
title: "Audible Dataset Data Cleaning Report"
subtitle: "BOS Analytical Unit | Technical Assessment"
author: "Tulsidai Singh"
date: today
format:
  docx:
    toc: true
    toc-depth: 2
    number-sections: true
    fig-width: 7
    fig-height: 4
    fig-dpi: 300
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(here)
source(here::here("R/00_config.R"))
library(ggplot2)
library(dplyr)
library(scales)
library(naniar)

df_raw   <- readRDS(file.path(p_out, "raw_loaded.rds"))
df_clean <- readRDS(file.path(p_out, "imputed.rds"))

# Colour palette
col_navy  <- "#1D2D44"
col_gold  <- "#C9A84C"
col_grey  <- "#8C8C8C"
col_light <- "#F5F5F5"

theme_audible <- function() {
  theme_minimal(base_size = 11, base_family = "sans") +
    theme(
      plot.title       = element_text(colour = col_navy, face = "bold", size = 13),
      plot.subtitle    = element_text(colour = col_grey, size = 10),
      plot.caption     = element_text(colour = col_grey, size = 8, hjust = 0),
      axis.title       = element_text(colour = col_navy, size = 10),
      axis.text        = element_text(colour = col_grey, size = 9),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(colour = "#EBEBEB"),
      plot.background  = element_rect(fill = "white", colour = NA),
      panel.background = element_rect(fill = "white", colour = NA),
      strip.text       = element_text(colour = col_navy, face = "bold")
    )
}
```

# Dataset Overview

The dataset used in this pipeline is the Audible audiobook dataset sourced from Kaggle
(CC0 Public Domain, collected by Snehangsude via web scraping from Audible.in).
It contains **87,489 records** across **8 variables** covering audiobook titles,
authors, narrators, duration, release dates, languages, star ratings, and prices in INR.
The data spans releases from December 1998 through November 2025.

| Variable | Description |
|---|---|
| `name` | Audiobook title |
| `author` | Author(s) |
| `narrator` | Narrator(s) |
| `time` | Duration as text string |
| `releasedate` | Release date (DD-MM-YY) |
| `language` | Language of the audiobook |
| `stars` | Rating string |
| `price` | Price in INR |

# Data Quality Diagnostics

## Missingness

No true NA values were present in the raw dataset. However, two variables contained
placeholder strings used in place of missing values: `stars` had 72,417 entries
(82.8%) recorded as `"Not rated yet"`, and `price` had 338 entries (0.4%) recorded
as `"Free"`. These were treated as structural missing and encoding issues respectively,
not as data errors.

```{r fig-missingness}
#| fig-cap: "Missingness by variable, placeholders recoded to NA for visualisation"

df_vis <- df_raw
df_vis$stars[df_vis$stars == "Not rated yet"] <- NA
df_vis$price[df_vis$price == "Free"] <- NA

miss_data <- miss_var_summary(df_vis) |>
  mutate(variable = reorder(variable, pct_miss),
         pct_miss = as.numeric(pct_miss))

ggplot(miss_data, aes(x = pct_miss, y = variable)) +
  geom_col(fill = col_navy, width = 0.6) +
  geom_text(aes(label = paste0(round(pct_miss, 1), "%")),
            hjust = -0.2, size = 3, colour = col_navy) +
  scale_x_continuous(limits = c(0, 100), labels = label_percent(scale = 1)) +
  labs(title = "Missingness by Variable",
       subtitle = "Placeholder strings recoded to NA for diagnosis",
       x = "% Missing", y = NULL,
       caption = "Source: audible_uncleaned.csv") +
  theme_audible()
```

## Column Types

All eight raw variables were ingested as character strings. Type coercion was required
for `price` (numeric), `releasedate` (date), `time` (numeric minutes), and `stars`
(split into numeric rating and integer count).

## Duplicates

No full-row duplicates were found. 1,523 records share the same `name` and `author`
but differ by narrator, language, release date, or price and confirmed as distinct
editions. No deduplication was applied.

# Cleaning Decisions

## Author and Narrator

Both fields contained a concatenated prefix (`Writtenby:`, `Narratedby:`) with no
spaces between words (e.g. `GeronimoStilton`). The prefix was stripped using regex
and word boundaries were inserted between camelCase joins. This was verified to work
correctly on Latin, accented, and Asian-script names. A known edge case exists where
compound lowercase particles (e.g. `de`, `van`) following an initial may not be
separated but it was documented as a limitation.

## Language

All language values were normalised to title case and underscores replaced with spaces
(`mandarin_chinese` → `Mandarin Chinese`). Only `English` and `Hindi` were already
title case; all other 34 languages were lowercase.

## Price

Comma separators were stripped from values like `"1,002.00"`. The 338 `"Free"` entries
were converted to `0` rather than `NA`, free audiobooks are a legitimate price point,
not missing data. Converted to numeric.

## Stars

The concatenated string `"4.5 out of 5 stars41 ratings"` was split into two columns:
`rating` (numeric, 0–5) and `rating_count` (integer). The 72,417 `"Not rated yet"`
entries were converted to `NA` for both columns. This is a structural missing error,
83% of the dataset is unrated, reflecting genuine absence of ratings rather than
a data collection failure.

## Time

Duration strings were parsed into `duration_mins` (total minutes, numeric). All
format variants were handled: `"2 hrs and 20 mins"`, `"10 hrs"`, `"1 hr and 1 min"`,
`"1 min"`, and `"Less than 1 minute"` (converted to 0).

## Release Date

Dates stored as `DD-MM-YY` were parsed to proper R date objects. A century cutoff
rule was applied: 2-digit years 95–99 map to 1995–1999; 00–94 map to 2000–2094.
This is justified by the dataset's documented range of 1998–2025. No NAs were
produced by the parse.

## Short Name Flagging

50 records with titles shorter than 3 characters were flagged via `short_name_flag`.
These are likely legitimate titles (single characters, numbers, Japanese kanji) and
were not removed.

## Outlier Flagging

IQR-based outlier detection (k = 1.5) was applied to `price` and `duration_mins`.
704 price outliers and 2,098 duration outliers were flagged. No records were dropped, but
flags are available for downstream analysis.

# Imputation

Only `rating` and `rating_count` contained missing values after cleaning (72,417 NAs,
82.8%). Imputation was not applied. With 83% of the column missing, any imputed value
would reflect the distribution of the rated minority, not the true population. The
missingness is structural, books without ratings on Audible simply have no rating.
Leaving these as `NA` is the honest and defensible choice.

All other variables had zero missing values after cleaning.

# Post-Cleaning Summary

```{r tbl-summary}
summary_data <- data.frame(
  Metric = c("Total records", "Total variables (after cleaning)",
             "Records with price = 0 (Free)", "Records flagged: short name",
             "Records flagged: price outlier", "Records flagged: duration outlier",
             "Missing: rating / rating_count", "Missing: all other variables"),
  Value = c("87,489", "14", "338 (0.4%)", "50 (0.1%)",
            "704 (0.8%)", "2,098 (2.4%)", "72,417 (82.8%)", "0")
)

knitr::kable(summary_data, col.names = c("Metric", "Value"),
             caption = "Post-cleaning dataset summary")
```

```{r fig-price}
#| fig-cap: "Distribution of audiobook price (INR) after cleaning"

q1_p  <- quantile(df_clean$price, 0.25)
q3_p  <- quantile(df_clean$price, 0.75)
iqr_p <- q3_p - q1_p
upper_p <- q3_p + 1.5 * iqr_p

ggplot(df_clean, aes(x = price)) +
  geom_histogram(bins = 60, fill = col_navy, colour = "white", linewidth = 0.2) +
  geom_vline(xintercept = upper_p, colour = col_gold, linetype = "dashed", linewidth = 0.8) +
  annotate("text", x = upper_p + 50, y = Inf, vjust = 2,
           label = paste0("IQR outlier\nthreshold: ", round(upper_p)),
           colour = col_gold, size = 3, hjust = 0) +
  scale_x_continuous(labels = label_comma()) +
  labs(title = "Price Distribution",
       subtitle = "Prices in Indian Rupees (INR) | Free entries converted to 0",
       x = "Price (INR)", y = "Count",
       caption = "Dashed line: IQR outlier threshold (k = 1.5)") +
  theme_audible()
```

```{r fig-duration}
#| fig-cap: "Distribution of audiobook duration (minutes) after cleaning"

q1_d  <- quantile(df_clean$duration_mins, 0.25)
q3_d  <- quantile(df_clean$duration_mins, 0.75)
iqr_d <- q3_d - q1_d
upper_d <- q3_d + 1.5 * iqr_d

ggplot(df_clean, aes(x = duration_mins)) +
  geom_histogram(bins = 60, fill = col_navy, colour = "white", linewidth = 0.2) +
  geom_vline(xintercept = upper_d, colour = col_gold, linetype = "dashed", linewidth = 0.8) +
  annotate("text", x = upper_d + 20, y = Inf, vjust = 2,
           label = paste0("IQR outlier\nthreshold: ", round(upper_d), " mins"),
           colour = col_gold, size = 3, hjust = 0) +
  scale_x_continuous(labels = label_comma()) +
  labs(title = "Duration Distribution",
       subtitle = "Parsed from text strings into total minutes",
       x = "Duration (minutes)", y = "Count",
       caption = "Dashed line: IQR outlier threshold (k = 1.5)") +
  theme_audible()
```

```{r fig-rating}
#| fig-cap: "Distribution of star ratings among rated audiobooks"

df_rated <- df_clean |> filter(!is.na(rating))

ggplot(df_rated, aes(x = factor(rating))) +
  geom_bar(fill = col_navy, width = 0.6) +
  geom_text(stat = "count", aes(label = after_stat(count)),
            vjust = -0.4, size = 3, colour = col_navy) +
  labs(title = "Star Rating Distribution",
       subtitle = paste0("Based on ", scales::comma(nrow(df_rated)),
                         " rated audiobooks (", scales::comma(nrow(df_clean) - nrow(df_rated)),
                         " unrated excluded)"),
       x = "Star Rating", y = "Count",
       caption = "Unrated entries (Not rated yet) excluded from this chart") +
  theme_audible()
```

```{r fig-language}
#| fig-cap: "Top 10 languages by number of audiobooks"

lang_counts <- df_clean |>
  count(language, sort = TRUE) |>
  slice_head(n = 10) |>
  mutate(language = reorder(language, n))

ggplot(lang_counts, aes(x = n, y = language)) +
  geom_col(fill = col_navy, width = 0.6) +
  geom_text(aes(label = scales::comma(n)),
            hjust = -0.2, size = 3, colour = col_navy) +
  scale_x_continuous(labels = label_comma(),
                     limits = c(0, max(lang_counts$n) * 1.15)) +
  labs(title = "Top 10 Languages",
       subtitle = "English dominates at 70.7% of the dataset",
       x = "Number of audiobooks", y = NULL,
       caption = "36 languages total in the dataset") +
  theme_audible()
```

# Limitations and Known Risks

- **Author/narrator word spacing** - the regex correctly handles standard camelCase
  but may miss spaces before lowercase particles (e.g. `de`, `van`, `al`) in
  non-English names. Affected records are a small minority and have been noted.

- **2-digit year inference** - the century cutoff rule is a judgment call. Any record
  with a release year of 95–99 has been mapped to the 1990s. If future data contains
  genuine 2095–2099 releases, this rule would need revisiting.

- **Price currency** - all prices are in Indian Rupees (INR), as this data was
  scraped from Audible.in. No currency conversion has been applied.

- **Structural missing in ratings** - 82.8% of rating values are `NA`. Any analysis
  using `rating` or `rating_count` applies only to the 17.2% of audiobooks that
  have been rated, and may not be representative of the full catalogue.

- **Web-scraped data** - the dataset was collected via scraping and may contain
  transcription errors, encoding artefacts, or records that were incomplete at the
  time of collection.